tstrait: a quantitative trait simulator for ancestral recombination graphs

### 2024-03-12 paper.tex

An important property of these population genetic simulation methods is that
they output ancestral recombination graphs (ARGs)
rather than sample genotypes.
ARGs encode the interwoven paths
of genetic inheritance arising from
% Explanation of citations: people might get offended if we don't cite
% Griffiths when first mentioning the ARG, but Hudson has just as
% much to do with the actual definition that we use in Wong 2023.
recombination~\citep{hudson1983properties,griffiths1997ancestral,wong2023general},
and contain rich detail about ancestral processes.
Recent breakthroughs in inferrence methods have made it possible
to estimate ARGs at biobank scale~\citep{kelleher2019,zhang2023},
and there is now intense interest in their practical
application~\citep{lewanski2023era}.
Statistical genetics has been a particular focus,
and ARG-based methods have been shown
to detect more ultra rare variants than conventional association testing
methods~\citep{zhang2023}; to
have better power to detect causal loci in
quantitative-trait locus mapping~\citep{link2023tree};
and to provide a sparse and efficient model of linkage disequilibrium
in GWAS and downstream applications~\citep{nowbandegani2023extremely}.

>>>>

An important property of these population genetic simulation methods is that
they output ancestral recombination graphs (ARGs)
rather than sample genotypes.
ARGs encode the interwoven paths
of genetic inheritance arising from
% Explanation of citations: people might get offended if we don't cite
% Griffiths when first mentioning the ARG, but Hudson has just as
% much to do with the actual definition that we use in Wong 2023.
recombination~\citep{hudson1983properties,griffiths1997ancestral,wong2023general},
and contain rich detail about ancestral processes.
Recent breakthroughs in inferrence methods have made it possible
to estimate ARGs at biobank scale~\citep{kelleher2019,zhang2023},
and there is now intense interest in their practical
application~\citep{lewanski2024era,brandt2024promise}.
Statistical genetics has been a particular focus,
and ARG-based methods have been shown
to detect more ultra rare variants than conventional association testing
methods~\citep{zhang2023}; to
have better power to detect causal loci in
quantitative-trait locus mapping~\citep{link2023tree};
and to provide a sparse and efficient model of linkage disequilibrium
in GWAS and downstream applications~\citep{nowbandegani2023extremely}.
